LME-MPM applied to quasi-brittle fracture.

### 2020-06-23 2020_EFM_MPM_Eigensoftening.tex
Updates from Overleaf

\begin{abstract}
  The objective of this work is to introduce an alternative
  technique to address the fracture process of brittle and
  quasi-brittle materials under the \acrfull{mpm} 
  framework. With this purpose the eigensoftening algorithm, developed
  originally for the \acrfull{otm} approximation scheme, is extended
  to the \acrshort{mpm} with the aim of presenting a suitable alternative
  to the existing fracture algorithms developed for the
  \acrshort{mpm}. The good fitting in the predictions made by the
  eigensoftening algorithm against both analytical and experimental
  results proofs the well performance of the method when challenging
  applications are to modeled.
\end{abstract}

@@

   \Grad{N_{Ip}}\ \Omega_p

@@

\subsection{$\epsilon$-neighbourhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
\PNA{The construction of the $\epsilon$-neighbourhood is a major issue of the proposed methodology.} Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms like Crossing Number \cite{Shimrat_1962} or Correct
Even-Odd \cite{Galetzka_et_al_2017}, \PNA{the proposed methodology adopted herein is linked to the shape function neighbourhood. In order to avoid the aforementioned error} prone
techniques, we endeavour to exploit the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global support of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore. Instead, a linked
list with the surrounding nodes for each node is defined at the initial
time. Also, each particle has to be associated with the
closest node instead with the belonging element as heretofore has been
done.
\MMP{Now lets define the concept of a particle \textit{tributary nodes} as the
list of nodes close to it among which the information transfers
occurs. Once the closest node are defined, the particle
posses automatically a list of \textit{tributary nodes}. This has two
immediate benefits, the first is that shape functions with global
support as \acrshort{lme} are easier to implement,
and the second is the increase of the efficiency of the $\epsilon$-neighbourhood
reconstruction. Regarding this last benefit, since the current
particle has a \textit{tributary nodes} list assigned to it, and
consequently each node has a linked list with a set of
\textit{tributary particles}\footnote{Notice that the list of the
  particles close to each node is unique for each node and is updated
  in each time step.} close to it, is possible to define efficiently a
search list with those particles close to the current particle.
In the end, particles out of the area of influence, Figure
\ref{fig:Failed-particles}, are pop out of the final
$\epsilon$-neighbourhood. A final remark about this new approach is
its ability to reconstruct easily a search list with the support of
the initial connectivity of the mesh. It provides a suitable basis for
an easy implementation of other interpolation techniques with minimum
coding effort. A detailed explanation of the proposed algorithm can be
found in \ref{sec:node-linked-method-1}.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\DM{The proposed approach to overcome the two main shortcomings of
  EighenMPM \cite{Zhang_EE_2020} has been evaluated with ``three
  benchmark tests''}.
The first one deals with the presence of stress
instabilities provoked by the spatial discretization, appearing even when the \acrshort{gimp} shape functions are employed. To overcome it, \acrshort{lme}
approximants are introduced as an alternative to the existing interpolation techniques. Its
performance in mitigating spurious stress oscillations under the \acrshort{mpm}~\cite{Wobbes2020,Molinos2020} framework helps to enhance
notoriously the quality of the results as it is observed in Section
\ref{sec:3.1}, where both interpolation techniques
are compared by carrying out an eigenerosion simulation. The second
limitation is concerned to its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}.\MMP{A proof
of it is exposed in Sections \ref{sec:3.2} \ref{sec:3.3}, where experimental results
are compared with experimental ones as well as eigensoftening
computations. In the first case comparing with a semicircular bending
test and in the second case with a drop-weight impact test.}

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation. The application
consists of a square plate of size\MMP{$H = 1\ [L] $} containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \MMP{\gls{E} = 1.06
$[M] [L]^{-1} [T]^{-2}$, the Poisson's ratio \gls{nu} = 0.333, and
critical energy-release rate \gls{Gf} = 0.0001 $[L]^{2} [T]^{-2}$. The absence of units is due to the current simulation will not be validated against experimental results, where
scale factor is relevant, but will be against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}.} In
\acrshort{mpm}, two different discretization are required. On one
hand, a cartesian grid of nodes is considered with a nodal spacing
value of 0.025. On the other hand, the plate will be modeled with a
initial layout of four particles per element and occupying the Gauss
quadrature positions.\\

\MMP{Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
\acrshort{lme}$_{\gamma = 3.5}$ solution has a reaction peak value of 0.03 $[M] [L]
[T]^{-2}$ for a imposed displacement 0.015 $[L]$ which agrees with the
analytical solution \cite{Pandolfi_2012}}. It also shows the presence
of wiggles in the reaction-displacements curve, in both loading and
post-failure stages, when linear interpolation technique is used. In
contrast, \acrshort{lme} simulation does not exhibit these spurious
oscillations and remains linear until the failure. About the strength,
linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. \PNA{Although, apparently, the
  differences in the obtained peaks are not extremely significant, it
  can be owing to the aforementioned hookean material
  employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive
  models will be employed, severe inaccuracies may appear in the
  stress field that could affect dramatically to the final 
  result. Focusing on the post failure behavior, it can be seen how
  \acrshort{lme} produces soft stress field evolution even after
  breaking.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: \PNA{once the
  energy is release dinamically by the crack, the waves propagate
  along the domain.}

@@

The second application to be considered is two semicircular
bending tests on Johnstone, which have been carried out by Lim {\it et
  al.} \cite{LIM_1993} and numerically reproduced by Wang {\it et
  al.}\cite{Wang_2020} under \acrshort{sph} framework within an
implicit algorithm. The aim of the test is to examine the predictive
capability of the eigensoftening algorithm under different kind of
mixed-mode loading conditions. This capability was previously within a
\acrshort{otm} implementation \cite{Navas_2018_ES}.

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test}. The experiment consist in two
semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm are supported by two rollers at a span of 47.5 mm and loaded by
another roller on its mid-span. Notches with a length of 16.6 mm are
created with different notch angles of 0$^º$ and 60$^º$ with respect
to the vertical axis to investigate the influences of the
notch angles on the peak load \ref{fig:plot-Angle-Forces-test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material properties of Johnstone are listed in Table \ref{tab:Johnstone-properties}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering
  \begin{tabular}[]{l l}
    \hline
    Young's modulus (\gls{E})   & 0.4\ $GPa$       \\
    Poisson ratio (\gls{nu})    & 0.25           \\
    Density (\gls{rho})         & 1.54\ $g/cm^3$ \\
    Tensile strength (\gls{ft}) & 0.6\ $MPa$       \\
    Critical opening displacement (\gls{Wc}) & 0.015\ $mm$ \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the tests, particles were uniformly distributed with an almost
constant separation of 1 mm between them, resulting in a total number of 4672
\acrshort{mpm} particles for generating the specimen. Additionally a
regular cartesian background grid of 1917 nodes with 1.5 mm of nodal distance
was generated. The velocities of the top boundary particles are $v_x$ = 0, $v_y$ = −
2 mm/s, whereas the velocities of the bottom boundary particles are
$v_y$ = 0 mm/s. To reduce dynamic effects due to a sudden loading, it
was preformed within a soft loading ramp. Finally, the model was
calibrated with a regularization parameter \gls{Ceps} = 1.7 and a
bandwidth value \gls{heps} = 15.

Figure \ref{fig:Figure-Angle-Forces-test-damage-pattern} shows the
comparison of mixed-mode fracture envelope between experiments and
\acrshort{mpm} simulations for different notch inclination angles
$\alpha$ = 0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$. The final
fracture pattern is a vertical straight line when $\alpha$ = 0$^{\circ}$  (pure mode I). 
When $\alpha$ > 0$^{\circ}$, the sample is
subjected to mixed-mode. For simulation cases 0$^{\circ}$ and 30$^{\circ}$, the fractures
initiate from the notch tip and then propagate towards upper loading
point. Experiments show that a few fractures tend to initiate behind
notch tip when the notch angle is beyond 50$^{\circ}$, figure
\ref{fig:Figure-Angle-Forces-test-damage-pattern} demonstrate that
eigensoftening combined with \acrshort{mpm} is able to reproduce this
behavior. Therefore, the simulation results predict successfully the
fracture envelope in experiments and demonstrate the capability of
the eigensoftening-\acrshort{mpm} model for capturing the fracture
development under various complex loading conditions. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[$\alpha$ = 0]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-0-damage-pattern.png}}
\subfigure[$\alpha$ = 30]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-30-damage-pattern.png}}
\subfigure[$\alpha$ = 60]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-60-damage-pattern.png}}
\caption{Crack pattern for different notch inclination angle, and
  comparison with the experimental results  \cite{LIM_1993}.}
\label{fig:Figure-Angle-Forces-test-damage-pattern}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, \ref{fig:plot-Angle-Forces-test} compare the peak load
between experiments and simulations for the notch inclination angles
0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$. This figure proofs how the
eigensoftening-\acrshort{mpm} simulation can well capture the
increasing trend of the peak load with the increasing notch angles in
the experiment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{./Figure-Angle-Forces-test}
  \caption{Numerical results compared with the experimental results \cite{LIM_1993}.}
  \label{fig:plot-Angle-Forces-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\begin{abstract}
  The objective of this work is to introduce an alternative
  technique to address the fracture process of brittle and
  quasi-brittle materials under the \acrfull{mpm} 
  framework. With this purpose the eigensoftening algorithm, developed
  originally for the \acrfull{otm} approximation scheme, is extended
  to the \acrshort{mpm} with the aim of presenting a suitable alternative
  to the existing fracture algorithms developed for \PNA{such a methodology}.
  %\acrshort{mpm}. 
  The good fitting in the predictions made by the
  eigensoftening algorithm against both analytical and experimental
  results proofs the \PNA{excellent} performance of the method when challenging
  applications are to modeled.
\end{abstract}

@@

   \Grad{N_{Ip}}\ \Omega_p,

@@

\subsection{$\epsilon$-neighbourhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
\PNA{The construction of the $\epsilon$-neighbourhood is a major issue of the proposed methodology.} Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms like Crossing Number \cite{Shimrat_1962} or Correct
Even-Odd \cite{Galetzka_et_al_2017}, \PNA{the proposed methodology adopted herein is linked to the shape function neighbourhood. In order to avoid the aforementioned error} prone
techniques, we endeavour to exploit the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global support of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore. Instead, a linked
list with the surrounding nodes for each node is defined at the initial
time. Also, each particle has to be associated with the
closest node instead of with the belonging element as heretofore has been
done.
\MMP{Now lets define the concept of a particle \textit{tributary nodes} as the
list of nodes close to it among which the information transfers
occurs. Once the closest nodes are defined, the particle
possesses automatically a list of \textit{tributary nodes}. This has two
immediate benefits: the first one is that shape functions with global
support as \acrshort{lme} are easier to be implemented,
and the second one is the increase of the efficiency of the $\epsilon$-neighbourhood
reconstruction. Regarding this last benefit, since the current
particle has a \textit{tributary nodes} list assigned to it, and
consequently each node has a linked list with a set of
\textit{tributary particles}\footnote{Notice that the list of the
  particles close to each node is unique for each node and is updated
  in each time step.} close to it, it is possible to define efficiently a
search list with those particles close to the current particle.
In the end, particles out of the area of influence, Figure
\ref{fig:Failed-particles}, are pop out of the final
$\epsilon$-neighbourhood. A final remark about this new approach is
its ability to reconstruct easily a search list with the support of
the initial connectivity of the mesh. It provides a suitable basis for
an easy implementation of other interpolation techniques with minimum
coding effort. A detailed explanation of the proposed algorithm can be
found in \ref{sec:node-linked-method-1}.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\DM{The proposed approach to overcome the two main shortcomings of
  EighenMPM \cite{Zhang_EE_2020} has been evaluated with ``three
  benchmark tests''}.
The first one deals with the presence of stress
instabilities provoked by the spatial discretization, appearing even when the \acrshort{gimp} shape functions are employed. To overcome it, \acrshort{lme}
approximants are introduced as an alternative to the existing interpolation techniques. Its
performance in mitigating spurious stress oscillations under the \acrshort{mpm}~\cite{Wobbes2020,Molinos2020} framework helps to enhance
notoriously the quality of the results as it is observed in Section
\ref{sec:3.1}, where both interpolation techniques
are compared by carrying out an eigenerosion simulation. The second
limitation is concerned to its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}.\MMP{A proof
of it is exposed in Sections \ref{sec:3.2} and \ref{sec:3.3}, where experimental results
are compared \PNA{with numerical ones, being the eigensoftening technique the one employed in this research for the
computations.} In the first case comparing with a semicircular bending
test and in the second case with a drop-weight impact test.}

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation. The application
consists of a square plate of size\MMP{$H = 1\ [L] $} containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \MMP{\gls{E} = 1.06
$[M] [L]^{-1} [T]^{-2}$, the Poisson's ratio \gls{nu} = 0.333, and
critical energy-release rate \gls{Gf} = 0.0001 $[L]^{2} [T]^{-2}$. The absence of units is due to the current simulation will not be tested against experimental results, where
scale factor is relevant, but will be validated against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}.} In
\acrshort{mpm}, two different discretizations are required. On the one
hand, a cartesian grid of nodes is considered with a nodal spacing
value of 0.025 $[L]$. On the other hand, the plate will be modeled with a
initial layout of four particles per element and occupying the Gauss
quadrature positions.\\

\MMP{Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows that the
\acrshort{lme}$_{\gamma = 3.5}$ solution has a reaction peak value of 0.03 $[M] [L]
[T]^{-2}$ for a imposed displacement 0.015 $[L]$ which agrees with the
analytical solution \cite{Pandolfi_2012}}. It also shows the presence
of wiggles in the reaction-displacements curve, in both loading and
post-failure stages, when linear interpolation technique is used. In
contrast, \acrshort{lme} simulation does not exhibit these spurious
oscillations and remains linear until the failure. About the strength,
linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. \PNA{Although, apparently, the
  differences in the obtained peaks are not extremely significant, it
  can be owing to the aforementioned hookean material
  employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive
  models will be employed, severe inaccuracies may appear in the
  stress field that could affect dramatically to the final 
  result. Focusing on the post failure behavior, it can be seen how
  \acrshort{lme} produces soft stress field evolution even after
  breaking.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: \PNA{once the
  energy is released dinamically by the crack, the waves propagate
  along the domain.}

@@

The second application to be considered \PNA{analyzes the bending behavior of semicircular specimens of Johnstone rocks, text which were experimentally} carried out by Lim {\it et
  al.} \cite{LIM_1993} and numerically reproduced by Wang {\it et
  al.}\cite{Wang_2020} under \acrshort{sph} framework within an
implicit algorithm. The aim of the test is to examine the predictive
capability of the eigensoftening algorithm under different kind of
mixed-mode loading conditions. This capability was assessed previously within a
\acrshort{otm} implementation \cite{Navas_2018_ES}.

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test}. The experiment \PNA{consist of loading
semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm, supported by two rollers at a span of 47.5 mm, by
another roller on its top mid-span. Notches with a length of 16.6 mm are
created with the different notch angles of 0$^º$, 30$^º$ and 60$^º$ with respect
to the vertical axis in order to investigate the influence of the
notch angle on the peak load \ref{fig:plot-Angle-Forces-test}.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material properties of Johnstone rock are listed in Table \ref{tab:Johnstone-properties}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering
  \begin{tabular}[]{l l}
    \hline
    Young's modulus (\gls{E})   & 0.4\ $GPa$       \\
    Poisson ratio (\gls{nu})    & 0.25           \\
    Density (\gls{rho})         & 1.54\ $g/cm^3$ \\
    Tensile strength (\gls{ft}) & 0.6\ $MPa$       \\
    Critical opening displacement (\gls{Wc}) & 0.015\ $mm$ \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the tests, particles were uniformly distributed with an almost
constant separation of 1 mm between them, resulting in a total number of 4672
\acrshort{mpm} particles for generating the specimen. Additionally, a
regular cartesian background grid of 1917 nodes with 1.5 mm of nodal distance
was generated. The velocities of the top boundary particles are $v_x$ = 0, $v_y$ = −
2 mm/s, whereas the velocities of the bottom boundary particles are
$v_y$ = 0 mm/s. To reduce dynamic effects due to a sudden loading, \PNA{a soft loading ramp has been employed as a loading curve.} Finally, the model was
calibrated with a regularization parameter \gls{Ceps} = 1.7 and a
bandwidth value \gls{heps} = 15.

Figure \ref{fig:Figure-Angle-Forces-test-damage-pattern} shows the
comparison of mixed-mode fracture envelope between experiments and
\acrshort{mpm} simulations for different notch inclination angles
$\alpha$ = 0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$. The final
fracture pattern is a vertical straight line when $\alpha$ = 0$^{\circ}$  (pure mode I). 
When $\alpha > 0^{\circ}$, the sample is
subjected to a mixed-mode pattern. For the cases of 0$^{\circ}$ and 30$^{\circ}$, cracks
initiate from the notch tip and then propagate towards upper loading
point. Experiments show that a few cracks tend to initiate behind
notch tip when the notch angle is beyond 50$^{\circ}$. Figure
\ref{fig:Figure-Angle-Forces-test-damage-pattern} demonstrates that
eigensoftening, combined with \acrshort{mpm}, is able to reproduce this
behavior. Therefore, the simulation results predict successfully the
fracture envelope in experiments and demonstrate the capability of
the eigensoftening-\acrshort{mpm} model for capturing the fracture
development under various complex loading conditions. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[$\alpha$ = 0]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-0-damage-pattern.png}}
\subfigure[$\alpha$ = 30]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-30-damage-pattern.png}}
\subfigure[$\alpha$ = 60]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-60-damage-pattern.png}}
\caption{Crack pattern for different notch inclination angle, and
  comparison with the experimental results  \cite{LIM_1993}.}
\label{fig:Figure-Angle-Forces-test-damage-pattern}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, \ref{fig:plot-Angle-Forces-test} compare the peak load
between experiments and simulations for the notch inclination angles
0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$. This figure proofs how the
eigensoftening-\acrshort{mpm} simulation can appropriately capture the
increasing trend of the peak load with the increasing notch angles in
the experiment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{./Figure-Angle-Forces-test}
  \caption{Numerical results compared with the experimental results \cite{LIM_1993}.}
  \label{fig:plot-Angle-Forces-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
